---
title: "Ntambi targeted Cleanup"
author: "Lainy VonBank"
date: "10/4/2024"
output: html_document
---

---

# Load packages
```{r}
if(!require(tidyverse)){install.packages("tidyverse")}
if(!require(dplyr)){install.packages("dplyr")}
if(!require(Cairo)){install.packages("Cairo")}
library(readxl)
getwd()
```

# Load data output from Agilent Quantitation software
```{r, message = F}
list.files()
mydata <- read_csv("quant_liver_pmol.csv")
Y <- mydata[-1,]
Y1 <- t(Y)
Y1 <- as.data.frame(Y1)
colnames(Y1) <- Y1[1,]
Y2 <- Y1[-1,]
Y3 <- type.convert(Y2,as.is = T)
```


# Blank subtraction 
Remove features where the abundance in an non-sample blank is higher than the average sample abundance
```{r}

Y3$avg <- rowMeans(Y3)
# Specify the name of the process blank column below after the subtraction
a <- as.name(colnames(select(Y3,matches("instd"))))
print(a)
# Copy this into the second part of the equation below. Make sure it is the PB

Y4 <- mutate(Y3, sub_height = `avg` - `2024_1024_instd_017.d`)
# Set NA values to 0
Y4$sub_height <- replace_na(Y4$sub_height, 0)
print(rownames(Y4[Y4$sub_height < 0,]))
# If any values are less than 0 in Y1$sub_height, that lipid(s) is noise or a contaminant and excluded
Y5 <- Y4[!Y4$sub_height < 0,]
```

# Remove extraneous columns and rename
```{r}

Y5 <- Y5 %>% select(-grep("instd", colnames(Y5)))
Y5 <- Y5 %>% select(-grep("avg", colnames(Y5)))
Y5 <- Y5 %>% select(-grep("sub_", colnames(Y5)))

colnames(Y5) <- gsub(".*plasma_", "", colnames(Y5))
colnames(Y5) <- gsub(".*liver_", "", colnames(Y5))
colnames(Y5) <- gsub("_0.*" , "", colnames(Y5))
colnames(Y5) <- gsub("__0.*" , "", colnames(Y5))
print(colnames(Y5))
```

# Transpose data
```{r}
Y5$lipid <- rownames(Y5)
Y5 <- Y5 %>% relocate(lipid)
Y5 <- sapply(Y5, as.character)
Y6 <- t(Y5)
Y6 <- as.data.frame(Y6)
colnames(Y6) <- Y6[1,]
Y6 <- Y6[-1,]
Y6 <- type.convert(Y6,as.is = T)
Y6$Extraction <- rownames(Y6)
Y6 <- Y6 %>% relocate(Extraction)

```

# Add sample information and produce initial output
```{r}
list.files()
sample_info <- read_csv("Samples.csv")
# Check for any other sample id discrepancies
Y6$Extraction <- as.numeric(Y6$Extraction)
Y7 <- merge(sample_info, Y6, by = "Extraction")

```
Verify data looks good and output initial merge file for further QC and analysis
```{r}
Y7[,5:ncol(Y7)] <- Y7[,5:ncol(Y7)]/Y7$Liver_mg


Y8 <- Y7[,-c(2,4)]
colnames(Y8) <- gsub(" Results","",colnames(Y8))

write_csv(Y8, "liver_cleanup_pmol_mgtissue.csv")
#write_csv(Y8, "plasma_cleanup_nM.csv")

rm(list = ls())
```


# Triglyceride pool saturation
```{r}
library(ggplot2)
library(ggpubr)
list.files()

Y9 <- read_csv("plasma_cleanup_nM.csv")

TG<- Y9[,c(1:2,grep("TG", colnames(Y9)))]
colnames(TG) <- gsub("NL.*","",colnames(TG))
colnames(TG) <- gsub("SIM.*","",colnames(TG))
colnames(TG) <- gsub(" \\[\\.*","",colnames(TG))

mdat2 <- reshape2::melt(TG, id = c(1:2))
mdat2$variable <- gsub(".*TG", "", mdat2$variable)
mdat2$variable <- gsub("[()]", "", mdat2$variable)

mdat3 <- str_split_fixed(mdat2$variable,":", 2)
mdat4 <- cbind(mdat2,mdat3)

names(mdat4)[names(mdat4) == '1'] <- 'length'
names(mdat4)[names(mdat4) == '2'] <- 'DB'


mdat5 <- aggregate(mdat4$value, by=list(Gen=mdat4$Genotype, DB=mdat4$DB), FUN=sum)
colnames(mdat5)[colnames(mdat5) == 'x'] <- 'value'

mdat5[,2] <- sapply(mdat5[,2], as.numeric)
mdat5$class[mdat5$DB == 0] <- "0"
mdat5$class[mdat5$DB == 1] <- "1"
mdat5$class[mdat5$DB == 2] <- "2"
mdat5$class[mdat5$DB == 3] <- "3"
mdat5$class[mdat5$DB == 4] <- "4"
mdat5$class[mdat5$DB == 5] <- "5"
mdat5$class[mdat5$DB > 5] <- "6+"



mdat5$Gen = factor(mdat5$Gen, levels = c("HET","GKO","TG5", "WT LL2","GKO LL2"), ordered = T)
mdat5$class = factor(mdat5$class, ordered = T)

p0 <- ggplot(mdat5, aes(x=Gen, y=value, fill=class)) + 
    geom_bar(position="fill", stat="identity")

p1 <- ggpar(p0, legend = "right", legend.title = "Number of double bonds", 
               ylab = "% Abundance", title = "Triglyceride Composition")
p2 <- p1 + theme_bw() 

CairoPDF(file = paste0("Plasma_composition_TG.pdf"), height = 8, width = 12,)
  print(p2)
rm(list = ls())
```

#Phospholipids
```{r}
Y9 <- read_csv("plasma_cleanup_nM.csv")

X <- Y9[,c(1:2,grep("P", colnames(Y9)))]
X <- X[,-grep("L", colnames(X))]

X1 <- reshape2::melt(X, id = c(1:2))

X1$variable <- gsub(".*C", "", X1$variable)
X1$variable <- gsub(".*E", "", X1$variable)
X1$variable <- gsub(".*I", "", X1$variable)
X1$variable <- gsub(".*S", "", X1$variable)
X1$variable <- gsub(".*G", "", X1$variable)
X1$variable <- gsub("[()]", "", X1$variable)
X1$variable <- gsub("OH", "", X1$variable)
X1$variable <- gsub("\\+", "", X1$variable)
X1$variable <- gsub("\\[]", "", X1$variable)
X1$variable <- gsub("a", "", X1$variable)
X1$variable <- gsub("b", "", X1$variable)
X1$variable <- gsub("c", "", X1$variable)
X1$variable <- gsub(" n6", "", X1$variable)
X1$variable <- gsub("O-", "", X1$variable)
X1$variable <- gsub("P-", "", X1$variable)
X1$variable <- gsub("\\\\","", X1$variable)
X1$variable <- gsub("/","_", X1$variable)
X1$variable <- gsub(" ","", X1$variable)

check <- unique(X1$variable)

X2 <- str_split_fixed(X1$variable, "_", 2)
X3 <- cbind(X1,X2)
X3_FA1 <- X3[,1:5]
X3_FA2 <- X3[,c(1:4,6)]


sn1 <- str_split_fixed(X3$`1`,":", 2)
sn2 <- str_split_fixed(X3$`2`,":", 2)
X6 <- cbind(X3,sn1,sn2)
X6[,7:10] <- sapply(X6[,7:10], as.numeric)
X6$sumacyl <- rowSums(X6[,c(7,9)], na.rm=T)
X6$sumdb <- rowSums(X6[,c(8,10)], na.rm=T)
X6$sumdb <- as.character(X6$sumdb)
X7 <- X6 %>% unite("sumCOMP",11:12, sep = ":", remove = F)
mdat5 <- aggregate(X7$value, by=list(Gen=X7$Genotype, DB=X7$sumdb), FUN=sum)
colnames(mdat5)[colnames(mdat5) == 'x'] <- 'value'

mdat5[,2] <- sapply(mdat5[,2], as.numeric)
mdat5$class[mdat5$DB == 0] <- "0"
mdat5$class[mdat5$DB == 1] <- "1"
mdat5$class[mdat5$DB == 2] <- "2"
mdat5$class[mdat5$DB == 3] <- "3"
mdat5$class[mdat5$DB == 4] <- "4"
mdat5$class[mdat5$DB == 5] <- "5"
mdat5$class[mdat5$DB > 5] <- "6+"



mdat5$Gen = factor(mdat5$Gen, levels = c("HET","GKO","TG5", "WT LL2","GKO LL2"), ordered = T)
mdat5$class = factor(mdat5$class, ordered = T)

p0 <- ggplot(mdat5, aes(x=Gen, y=value, fill=class)) + 
    geom_bar(position="fill", stat="identity")

p1 <- ggpar(p0, legend = "right", legend.title = "Number of double bonds", 
               ylab = "% Abundance", title = "Phospholipid Composition")
p2 <- p1 + theme_bw() 

CairoPDF(file = paste0("Plasma_Composition_PL.pdf"), height = 8, width = 12,)
  print(p2)
rm(list = ls())
```




# DAGs acyl chains
```{r}
list.files()
Y9 <- read_csv("liver_cleanup_pmol_mgtissue.csv")
X <- Y9[,c(1:2,grep("DG", colnames(Y9)))]

X1 <- reshape2::melt(X, id = c(1:2))

X1$variable <- gsub(".*G", "", X1$variable)
X1$variable <- gsub("[()]", "", X1$variable)

check <- unique(X1$variable)
print(check)

X2 <- str_split_fixed(X1$variable, "_", 2)
X3 <- cbind(X1,X2)
X3_FA1 <- X3[,1:5]
X3_FA2 <- X3[,c(1:4,6)]

names(X3_FA1)[5] <- "FA"
names(X3_FA2)[5] <- "FA"
mdat4 <- bind_rows(X3_FA1, X3_FA2)
mdat4$FA
chain <- unique(mdat4$FA) 

chain_keep <- chain
#chain_keep <- c("16:0","16:1","18:0","18:1")
X4 <- mdat4[mdat4$FA %in% chain_keep,]


Xagg <- aggregate(X4$value, by=list(Name=X4$Extraction, Genotype=X4$Genotype, FA=X4$FA), FUN=sum)
colnames(Xagg)[colnames(Xagg) == 'x'] <- 'value'

X5 <-  subset(Xagg,Genotype %in% c("HET","GKO"))
X5$Genotype <- factor(X5$Genotype,levels = c("HET","GKO"), ordered = T)
a = "FA"
b = "value"
c = "Genotype"

library(ggplot2)
library(ggpubr)
library(rstatix)
library(ggbreak)
p0 <- ggbarplot(X5, x = a, y = b,
            add = c("mean_se"), color = c,fill=c, 
            add.params = list(width = 0.35, binwidth = 0.07),
            alpha = 0.5, position = position_dodge(0.8), size = .5)  +
   scale_y_cut(breaks=c(2500), which=c(1, 2), scales=c(1, 2)) +
  #coord_cartesian(ylim=c(0,1.1*max(X5$value, na.rm = T)))
scale_y_continuous(limits = c(0.00, 1.1*max(X5$value, na.rm = T)), expand = c(0,0) ) 

p1 <- ggpar(p0, palette = c("red","blue"),legend = "right", title = "Diglyceride Acyl Chains",legend.title = c, xlab = F, ylab = "Acyl chain sum") +
  theme(text = element_text(size=12),
        axis.text.x = element_text(angle=45, hjust=1, size = 12)) 

stat.test <- X5 %>% 
                  group_by(FA) %>% 
                  pairwise_t_test(value ~ Genotype) %>% 
                  adjust_pvalue(method = "none") %>%
                  add_significance("p") 
                 
  stat.test$p.adj.signif <- stat.test$p.signif 
  
   stat.test <- stat.test %>%  add_xy_position(fun="mean_se",x=a)
  
  p3 <- p1 + stat_pvalue_manual(
      stat.test, label = "p.adj.signif", hide.ns = T,tip.length = 0.008)
 print(p3)

CairoPDF(file = paste0("Liver_DG_Chains.pdf"), height = 8, width = 12,)
  print(p3)


write_csv(stat.test,"Liver_DG_Chains_stats.csv")
rm(list = ls())
```